---
title: "Portfolio 1"
author: 'Søren Orm Hansen'
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pacman)

p_load(tidyverse, sjPlot, dplyr, nlme, ggplot2, pastecs)

```


```{r}
sleep <- read.csv('sleepstudy.csv')
```

###1.a: Get the data from one participant, e.g. using subset(). Make a linear regression for reaction time as a function of days of sleep deprivation, e.g. using lm(). Report the F-statistics.

```{r}
df <- subset(sleep, Subject == 352)

```

```{r}
regr1 <- lm(Reaction ~ Days, df)

summary(regr1)
```
Report f-stats:
The 



###1.b: How many degrees of freedom does the relevant F-distribution have?
One for the model and 8 for the residuals


###1.c: At which F-value does a regression with this distribution become statistically significant (p<0.05)? 

```{r}
f_val <- qf(0.95, 1, 8)

f_val
```

With a f-distribution 1 and 8 degrees of freedom the regression becomes


###1.d: Make a plot of the F-distribution.

```{r}
dist_f(f=0, deg.f1 = 1, deg.f2 = 8, xmax=7)
```


##2.a: Find the coefficients (slope and intercept) for the regression for reaction time as a function of days of sleep deprivation (a hint for the solution: use group_by() in tidyverse or this function here: https://stat.ethz.ch/R-manual/R-devel/library/nlme/html/lmList.html, hint2: pool=FALSE)

```{r}
sleep$Subject <- as.factor(sleep$Subject )

sleep <- group_by(sleep, Subject)

```

```{r}
fit <- lmList(Reaction ~ Days | Subject, data = sleep, pool = FALSE)

fit
```


##2.b: Combine both scatter plot and regression line in the same figure. You may also include all participants in one plot.

```{r}
ggplot(sleep, aes(Days, Reaction, color = Subject)) +
  geom_point() + 
  stat_smooth(method = lm, geom='line', alpha=0.5, se=FALSE)
  theme_minimal()



```


##2.c: Collect and report the inferential statistics for each participant in a table using t-statistics, including t-value, df and p-value.

```{r}
sum <- summary(fit)

df_coef <- data.frame(sum[["coefficients"]])

df_coef$ID <- row.names(df_coef)

inf_stats <- data.frame(df_coef$ID,df_coef$t.value.Days, df_coef$Pr...t...Days)

colnames(inf_stats)[1] <- "ID"
colnames(inf_stats)[2] <- "t-value_Days"
colnames(inf_stats)[3] <- "p-value_Days"

inf_stats$df_model <- 1
inf_stats$df_res <- 8

inf_stats
```


##2.d: How many individual participants display a statistically significant effect of sleep deprivation (p-values uncorrected for mulitple comparisons)?

All 18 of the participants are statistically significant with a p-value < 0.05


##3.a: Use the slopes you found for each participant in exercise 2 as a new dataset. Test the hypothesis that the slopes are larger than zero against the null-hypothesis that the slopes are zero (i.e. no differences in response time exist as a function of time).

```{r}
coefs <- coef(fit)

coefs <- select(coefs, Days)

coefs$null <- 0

t.test(coefs$Days, alternative = 'greater')

```








##3.b: Justify your use of test statistics.
```{r}
hist(coefs$Days)

round(stat.desc(coefs$Days, norm = T, basic = F), digits = 4)

```


We used a one sample t-test to test whether the slopes are signigicantly larger that 0.
The data is normally distributed (the shape of the histogram, the Shapiro Wilks test for normality, and the skew.2SE and kurt.2SE indicates normality).


##3.c: Report inferential statistics.

The t-test found that the slopes were singnificantly larger than 0, t(17) = 6.7715, SEM = 10.50, p < .001. 

##3.d: Make a plot with the mean reaction time and standard errror bars for each day across participants and plot the averaged regression line in the same figure.

```{r}
reg_line <- coef(lm(Reaction ~ Days, data = sleepstudy))

ggplot(sleepstudy, aes(x = Days, y = Reaction, fill = Days)) +
  scale_color_gradientn(colours = rainbow(4)) +
  geom_boxplot(width=0.5) +
  stat_summary(fun.y = mean, geom = "point", shape = 16, color="red") +
  stat_summary(fun.data = mean_se, geom = "errorbar", color = 'black', width = 0.1) +
  geom_abline(intercept = reg_line[1], slope = reg_line[2]) +
  ggtitle("Reaction time by days w/out sleep") +
  labs(x = "Days", y = "Reaction time /ms") +
  theme_minimal()

```
