---
title: "portfolio 1 experimental methods 2"
author: 'Astrid Olling, Daniel Christensen, Karoline Klitgaard, Pernille Jensen, Søren
  Orm '
date: "3/2/2020"
output:
  pdf_document: default
  html_document: default
---

Setting up knitr and loading the relevant packages:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
p_load(tidyverse, sjPlot, dplyr, nlme, ggplot2, pastecs)
```

Loading the sleep data to a data frame:
```{r}
sleepstudy <- read.csv("sleepstudy.csv")
```

##Tasks ####1. Plot the data

1.a: Get the data from one participant:  

```{r}
#We choose one random participant from the original sleepstudy dataframe
our_guy <- subset(sleepstudy, sleepstudy$Subject == 352)
```

Make a linear regression for reaction time as a function of days of sleep deprivation:
```{r}
#linear regression on reaction time by days of sleep deprivation on our random participant

summary(lm(Reaction ~ Days, our_guy))
```

Report the F-statistics:

An increase in number of days with sleep deprivation was found to significantly predict an increase in reaction time at F(1,8) = 23.32, p > 0.01


1.b: How many degrees of freedom does the relevant F-distribution have?

The degrees of freedom for the model is 1 and the degrees of freedom for the residuals is 8


1.c: At which F-value does a regression with this distribution become statistically significant (p<0.05)? 
```{r}

#we calculate the cut-off value at 0.95
qf(0.95, 1, 8)
```
A regression with this distribution becomes statistically significant at F-value 5.317655

1.d: Make a plot of the F-distribution.
```{r}
dist_f(f=0, deg.f1 = 1, deg.f2 = 8,xmax=10)
```

####2.For all participants in the experiment

2.a: Find the coefficients (slope and intercept) for the regression for reaction time as a function of days of sleep deprivation

```{r}

#We convert the subject values to factors
sleepstudy$Subject <- as.factor(sleepstudy$Subject)

#Slopes and intercepts for each participants
list <- lmList(Reaction ~ Days | Subject, data = sleepstudy, pool = FALSE)

list

```


2.b: Combine both scatter plot and regression line in the same figure. You may also include all participants in one plot.

```{r}

#group by subject
grouped_pers <- group_by(sleepstudy, Subject)

#creating a plot with all participants
ggplot(grouped_pers, aes(Days, Reaction, color = Subject)) +
  geom_point() + 
  stat_smooth(method = lm, geom='line', alpha=0.5, se=FALSE)+
  theme_minimal()+
    labs(title = "Reaction time/days of sleep deprivation by participants", x = "Days of sleep deprivation", y = "Reaction time (ms)")

```

2.c: Collect and report the inferential statistics for each participant in a table using t-statistics, including t-value, df and p-value.

```{r}
#Saving a summary of the output from the linear model:
listsummary <- summary(list)

#Creating a data frame from the output:
final <- data.frame(listsummary[["coefficients"]])

#with relevant columns
final <- data.frame(row.names(final), final$t.value.Days, final$Pr...t...Days)

#Adding columns with degrees of freedom:
final$df_model <- 1
final$df_res <- 8

#Changing the column names:
colnames(final)[1] <- "ID"
colnames(final)[2] <- "t values"
colnames(final)[3] <- "p values"
```

2.d: How many individual participants display a statistically significant effect of sleep deprivation (p-values uncorrected for mulitple comparisons)?

An increase in number of days with sleep deprivation is found to significanty predict an increase in reaction time for all participants except one. For subject (find the number) an increase in days with sleep deprivation is instead found to predict a decrease in reaction time. For all participants the slopes are significant at p > 0.05. 

####3. Across participants: 3.a: Use the slopes you found for each participant in exercise 2 as a new dataset. Test the hypothesis that the slopes are larger than zero against the null-hypothesis that the slopes are zero 

```{r}
#dataframe with only the coefficients
all_coefs <- coef(list, data = sleepstudy)

#t-test for testing the hypothesis
t.test(all_coefs$Days, alternative = "greater", paired = FALSE)

```

3.b: Justify your use of test statistics.

Checking if the data is normally distributed:
```{r}
#numeric inspection of data distribution
round(stat.desc(all_coefs$Days, norm = TRUE, basic = FALSE), digits=2)

#visual inspection of data distribution
hist(all_coefs$Days)
```

We conducted a one-tailed t-test testing the alternative hypothesis against the null hypothesis.
This test can be used because the data is sufficiently normally distributed: the shape of the histogram, the Shapiro Wilks test for normality, and the skew.2SE and kurt.2SE indicates normality. It should however be noticed that a bigger sample size would make the test more reliable.

3.c: Report inferential statistics.

The t-test is significant at t(17) = 6.7715, p-value = 1.632e-06, which leaves us to reject the null hypothesis that the sample mean is equal to zero.

3.d: Make a plot with the mean reaction time and standard errror bars for each day across participants and plot the averaged regression line in the same figure.
```{r}
#days as factor
sleepstudy$Days <- as.factor(sleepstudy$Days)

#boxplot for mean reaction time by days
reg_line <- coef(lm(Reaction ~ Days, data = sleepstudy))

ggplot(sleepstudy, aes(x = Days, y = Reaction)) +
  geom_boxplot(width=0.5, color = 'darkblue', fill = 'lightblue') +
  stat_summary(fun.y = mean, geom = "point", shape = 16, color="red") +
  stat_summary(fun.data = mean_se, geom = "errorbar", color = 'black', width = 0.1) +
  geom_abline(intercept = reg_line[1], slope = reg_line[2]) +
  ggtitle("Reaction time by days w/out sleep") +
  labs(x = "Days", y = "Reaction time /ms") +
  theme_minimal()

```


####Voluntory bonus task Add 10% white/uniform noise to the Reaction data. What happens to the single participant coefficients statistics and to the group effects?

```{r}
sleepstudy$noise_reaction = sleepstudy$Reaction+sleepstudy$Reaction*runif(10, min = -0.05, max = 0.05)

sleepstudy$Subject <- as.factor(sleepstudy$Subject)

sleepstudy$Days <- as.numeric(sleepstudy$Days)

list_noise <- lmList(noise_reaction ~ Days | Subject, data = sleepstudy, pool = FALSE)
summary(list_noise)

#dataframe with only the coefficients
all_coefs_noise <- coef(list_noise, data = sleepstudy)

#t-test for testing the hypothesis
t.test(all_coefs_noise$Days, alternative = "greater", paired = FALSE)

reg_line2 <- coef(lm(Reaction ~ Days, data = sleepstudy))

sleepstudy$Days <- as.factor(sleepstudy$Days)

ggplot(sleepstudy, aes(x = Days, y = noise_reaction)) +
  geom_boxplot(width=0.5, color = 'darkblue', fill = 'lightblue') +
  stat_summary(fun.y = mean, geom = "point", shape = 16, color="red") +
  stat_summary(fun.data = mean_se, geom = "errorbar", color = 'black', width = 0.1) +
  geom_abline(intercept = reg_line2[1], slope = reg_line2[2]) +
  ggtitle("Reaction time by days w/out sleep") +
  labs(x = "Days", y = "Reaction time /ms") +
  theme_minimal()



```

With the 10% white noise, the coefficients change slightly (see the table) and they are still significant.





